---
title: "Reading_Preparaing files"
author: "Luis Rincones"
date: "7/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
Code  to read, and load the files
For the prepare activities
```{r libraries, echo=FALSE}
library("readr")
library("tools")
library("fs")
library("dplyr")
library("tidyverse")
library("stringr")
library("janitor")
```
Get the files from the folder.
Make a list of the files to read (files_list)
Make a list of the files names (files_names)
```{r reading.], echo=FALSE}
path_files <- "~/Documents/Case_2/Fitabase_Data"
files_list <- list.files(path = path_files,recursive = F , 
                       all.files = FALSE, full.names = TRUE, pattern = "*.csv*")

files_names <- path_file(files_list)

```
Read and load each file
I could have decided before hand and work with less files but need to justify

Tengo que solucionar leer usando file path por ahora cargue los archivos
los puse todos en la raiz del proyecto.
Luego como poner el nombre automatico


```{r reading_files, echo=FALSE}
path = path_file
# name without the extension
file_name_1 <- file_path_sans_ext(files_names)
# just the string before first "_"
# files names to use for dataframe names
file_name_1 <- str_extract(file_name_1, "[^_]+")
# Challenge how to read the file in df with the name
# 
dailyActivity <- read.csv(files_names[1])
dailyCalories <- read.csv(files_names[2])
dailyIntensities <- read.csv(files_names[3])
dailySteps <- read.csv(files_names[4])
heartrate <- read.csv(files_names[5])
hourlyCalories <- read.csv(files_names[6])
hourlyIntensities <- read.csv(files_names[7])
hourlySteps <- read.csv(files_names[8])
minuteCaloriesNarrow <- read.csv(files_names[9])
minuteCaloriesWide <- read.csv(files_names[10])
minuteIntensitiesNarrow <- read.csv(files_names[11])
minuteIntensitiesWide <- read.csv(files_names[12])
minuteMETsNarrow <- read.csv(files_names[13])
minuteSleep <- read.csv(files_names[14])
minuteStepsNarrow <- read.csv(files_names[15])
minuteStepsWide <- read.csv(files_names[16])
sleepDay <- read.csv(files_names[17])
weightLogInfo <- read.csv(files_names[18])


```
Ya tengo los diferentes dataframes ahora que hago 

Prepare adataframe df_ID_file with 3 variables; IDs for the number of records in the file, Unique_IDs for the number of unique IDs in the file and File for the file name.
This is used to review the population in the file vs the universe of thirty persons indicated in the metadata. This will assist in how to use the files in relations with other files. 

Next 3 chunks one for each variable

```{r verificar_usuarios, echo=FALSE}
columnas <- c("IDs", "Unique_IDs", "File")
df_ID_file <- data.frame(matrix(nrow = 18, ncol = length(columnas)))
colnames(df_ID_file) <- columnas
df_ID_file$File[1] <- "dailyActivity"
df_ID_file$File[2] <- "dailyCalories"
df_ID_file$File[3] <- "dailyIntensities"
df_ID_file$File[4] <- "dailySteps"
df_ID_file$File[5] <- "heartrate"
df_ID_file$File[6] <- "hourlyCalories"
df_ID_file$File[7] <- "hourlyIntensities"
df_ID_file$File[8] <- "hourlySteps"
df_ID_file$File[9] <- "minuteCaloriesNarrow"
df_ID_file$File[10] <- "minuteCaloriesWide"
df_ID_file$File[11] <- "minuteIntensitiesNarrow"
df_ID_file$File[12] <- "minuteIntensitiesWide"
df_ID_file$File[13] <- "minuteMETsNarrow"
df_ID_file$File[14] <- "minuteSleep"
df_ID_file$File[15] <- "minuteStepsNarrow"
df_ID_file$File[16] <- "minuteStepsWide"
df_ID_file$File[17] <- "sleepDay"
df_ID_file$File[18] <- "weightLogInfo"

```
usuarios unicos

```{r unicos, echo=FALSE}

df_ID_file$IDs[1] <- nrow(dailyActivity)
df_ID_file$IDs[2] <- nrow(dailyCalories)
df_ID_file$IDs[3] <- nrow(dailyIntensities)
df_ID_file$IDs[4] <- nrow(dailySteps)
df_ID_file$IDs[5] <- nrow(heartrate)
df_ID_file$IDs[6] <- nrow(hourlyCalories)
df_ID_file$IDs[7] <- nrow(hourlyIntensities)
df_ID_file$IDs[8] <- nrow(hourlySteps)
df_ID_file$IDs[9] <- nrow(minuteCaloriesNarrow)
df_ID_file$IDs[10] <- nrow(minuteCaloriesWide)
df_ID_file$IDs[11] <- nrow(minuteIntensitiesNarrow)
df_ID_file$IDs[12] <- nrow(minuteIntensitiesWide)
df_ID_file$IDs[13] <- nrow(minuteMETsNarrow)
df_ID_file$IDs[14] <- nrow(minuteSleep)
df_ID_file$IDs[15] <- nrow(minuteStepsNarrow)
df_ID_file$IDs[16] <- nrow(minuteStepsWide)
df_ID_file$IDs[17] <- nrow(sleepDay)
df_ID_file$IDs[18] <- nrow(weightLogInfo)

```



```{r unicos, echo=FALSE}

df_ID_file$Unique_IDs[1] <- length(unique(dailyActivity$Id))
df_ID_file$Unique_IDs[2] <- length(unique(dailyCalories$Id))
df_ID_file$Unique_IDs[3] <- length(unique(dailyIntensities$Id))
df_ID_file$Unique_IDs[4] <- length(unique(dailySteps$Id))
df_ID_file$Unique_IDs[5] <- length(unique(heartrate$Id))
df_ID_file$Unique_IDs[6] <- length(unique(hourlyCalories$Id))
df_ID_file$Unique_IDs[7] <- length(unique(hourlyIntensities$Id))
df_ID_file$Unique_IDs[8] <- length(unique(hourlySteps$Id))
df_ID_file$Unique_IDs[9] <- length(unique(minuteCaloriesNarrow$Id))
df_ID_file$Unique_IDs[10] <- length(unique(minuteCaloriesWide$Id))
df_ID_file$Unique_IDs[11] <- length(unique(minuteIntensitiesNarrow$Id))
df_ID_file$Unique_IDs[12] <- length(unique(minuteIntensitiesWide$Id))
df_ID_file$Unique_IDs[13] <- length(unique(minuteMETsNarrow$Id))
df_ID_file$Unique_IDs[14] <- length(unique(minuteSleep$Id))
df_ID_file$Unique_IDs[15] <- length(unique(minuteStepsNarrow$Id))
df_ID_file$Unique_IDs[16] <- length(unique(minuteStepsWide$Id))
df_ID_file$Unique_IDs[17] <- length(unique(sleepDay$Id))
df_ID_file$Unique_IDs[18] <- length(unique(weightLogInfo$Id))

```

```{r files_numbers, echo=FALSE}

tibble(df_ID_file)
```
FINDINGS
This files have too few records

heartrate --> 14 but 2483658 records --> 177404 records per user
weightLogInfo --> 8 with 67 records --> 8 records per user

For sleep files 
minuteSleep --> 24 with 188521 records  --> 7855 records per user
sleepDay --> 24 with 413 records --> 17 records per user
checking with 
length(unique(as.Date(minuteSleep$date))) me dio 15
length(unique(as.Date(sleepDay$SleepDay))) me dio 14
ambos se pueden usar el sleepDay es mas compacto y facil de trabajar
se analizara la informacion por dia

weight log 
length(unique(as.Date(weightLogInfo$Date))) me dio 14
heartrate
length(unique(as.Date(heartrate$Time))) me dio 14

Amos archivos se pueden usar para complementar el analisis basico

All the rest have 33 unique users how ever the metadata indicate 30 users is the sample
may need to check users with lower number of records


```{r records_user, echo=FALSE}
rec_user <- df_ID_file$IDs / df_ID_file$Unique_IDs

unique(minuteSleep$date)

```


El usuario promedio en hourlyCalories tendria 649 lecturas peoro el usuario
4057192912 tiene 88
busquemos la distribucion pero antes revisemos si hace falta o no
Que impacto tiene esa falla de 33 vs 30? y los que son pocos

en el caso pesos la mayoria lo reporto manualmente
> weightLogInfo %>% count(IsManualReport)
  IsManualReport  n
1          False 26
2           True 41

```{r unique_number, echo=FALSE}
usuarios <- dailyCalories %>% count(Id)
# Id 4057192912 tiene 4 veamos otro file con la misma Id


```

Sršen tells you that this data set might have some limitations, and encourages you to consider adding another data to help address those limitations as you begin to work more with this data.

Cual data esta disponible para complementar pero primero que tengo

```{r combinar_merge, echo=FALSE}
metnarrow_calories <- merge(minuteMETsNarrow, minuteCaloriesNarrow, all=TRUE)
n_distinct(minuteMETsNarrow$Id)

day_steps_sleep <- merge(dailySteps, sleepDay, all = TRUE)
n_distinct(dailySteps$Id)


# graficar para ver tendencia
ggplot(data = metnarrow_calories) +
        geom_point(aes(x=METs, y=Calories), color = "blue", alpha = 0.5) +
        labs(title = "Relation between METs and Calories", x = "METs", y = "Calories")
# No hay relacion al menos visual

ggplot(data = day_steps_sleep) +
        geom_point(aes(x=StepTotal, y=TotalMinutesAsleep), color = "red", alpha = 0.5) +
        labs(title = "Relation between Steps and Sleep", x = "Steps", y = "Minutes_Sleep")

```


```{r merge-day-calories, echo=FALSE}
# will not use for final analysis
dailyact2 <- merge(dailyIntensities, dailyCalories, by="Id")

```


dailyActivity, if the records that add to full day and records that do not. Have the same behaviour

FALSE has a more disperse correlation between totalsteps and calories
The correlations values are 0.733 when the records have a full day
and 0.46 when the records do not have a full day

```{r daylyAct_records, echo=FALSE}
copy_dailyActivity <- dailyActivity


copy_dailyActivity <-    copy_dailyActivity %>% 
  mutate(total = VeryActiveMinutes + FairlyActiveMinutes +
                        LightlyActiveMinutes + SedentaryMinutes)

copy_dailyActivity <- copy_dailyActivity %>% 
  mutate(valido = total == 1440)
   

copy_dailyActivity_TRUE <- subset(copy_dailyActivity, valido == TRUE)
copy_dailyActivity_FALSE <- subset(copy_dailyActivity, valido == FALSE)


ggplot(data = copy_dailyActivity_TRUE) +
        geom_point(aes(x=TotalSteps, y=Calories), color = "blue", alpha = 0.5) +
        labs(title = "Relation between Steps and Calories", x = "Steps", y = "Calories")

ggplot(data = copy_dailyActivity_FALSE) +
        geom_point(aes(x=TotalSteps, y=Calories), color = "red", alpha = 0.5) +
        labs(title = "Relation between Steps and Calories", x = "Steps", y = "Calories")

cor(copy_dailyActivity_TRUE$TotalSteps, copy_dailyActivity_TRUE$Calories)

cor(copy_dailyActivity_FALSE$TotalSteps, copy_dailyActivity_FALSE$Calories)


```

Merge copy_dailyActivity_TRUE with heartrate
Merge copy_dailyActivity_TRUE with sleepDay
Merge copy_dailyActivity_TRUE with weightLogInfo

```{r merge_3, echo=FALSE}
dailyact_true_heart <- merge(copy_dailyActivity_TRUE, heartrate, by="Id")
dailyact_true_sleep <- merge(copy_dailyActivity_TRUE, sleepDay, by="Id")
dailyact_true_weight <- merge(copy_dailyActivity_TRUE, weightLogInfo, by="Id")
dailyact_true_METs<- merge(copy_dailyActivity_TRUE, minuteMETsNarrow, by="Id")

minuteMETsNarrow
# try cor
cor(dailyact_true_heart$TotalSteps, dailyact_true_heart$Value)
#  0.05 low correlation
cor(dailyact_true_sleep$TotalSteps, dailyact_true_sleep$TotalMinutesAsleep)
# negative correlation  -0.2549887
cor(dailyact_true_weight$TotalSteps, dailyact_true_weight$WeightKg)
# low correlation  0.1238845
cor(dailyact_true_METs$TotalSteps, dailyact_true_METs$METs)
# low correlation  0.1393237
cor(dailyact_true_METs$TotalSteps, dailyact_true_METs$Calories)
# High correlation  0.7323676


```


