---
title: "Analyze"
author: "Luis Rincones"
date: "7/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
Code for the process phase, assume the data frames are available in the global environment of the Project.

Libraries load

```{r libraries, echo=FALSE}
library("readr")
library("tools")
library("fs")
library("dplyr")
library("tidyverse")
library("stringr")
library("janitor")
```
##Analyze
*Guiding questions* 
* How should you organize your data to perform analysis on it?<br> 
* Has your data been properly formatted?<br>
* What surprises did you discover in the data?<br> 
* What trends or relationships did you find in the data?<br> 
* How will these insights help answer your business questions?<br>

*Key tasks*
1. Aggregate your data so it’s useful and accessible<br> 
2. Organize and format your data<br> 
3. Perform calculations<br> 
4. Identify trends and relationships<br>

*Deliverable A summary of your analysis*

To get started in R
Here is a sample starter script that can help you do the following:
1.Begin importing your data. DONE
2.Explore your data, gathering some summary statistics DONE
3.Clean and transform your data to prepare for analysis DONE
4.Create some initial exploratory visualizations DONE



\newpage
## lo que tenia antes


**Ya tengo los diferentes dataframes ahora que hago?**<br>

Archivos combinados usando merge<br>
* metnarrow_calories<br>
* day_steps_sleep<br>
* dailyact_true_heart<br>
* dailyact_true_sleep<br>
* dailyact_true_weight<br> 
* dailyact_true_METs<br>

Merging files and subsetting sets

```{r combinar_merge, echo=FALSE}
metnarrow_calories <- merge(minuteMETsNarrow, minuteCaloriesNarrow, all=TRUE)
n_distinct(minuteMETsNarrow$Id)

day_steps_sleep <- merge(dailySteps, sleepDay, all = TRUE)
n_distinct(dailySteps$Id)

dailyact_true_heart <- merge(copy_dailyActivity_TRUE, heartrate, by="Id")
n_distinct(dailyact_true_heart$Id)
dailyact_true_sleep <- merge(copy_dailyActivity_TRUE, sleepDay, by="Id")
n_distinct(dailyact_true_sleep$Id)
dailyact_true_weight <- merge(copy_dailyActivity_TRUE, weightLogInfo, by="Id")
n_distinct(dailyact_true_weight$Id)
dailyact_true_METs<- merge(copy_dailyActivity_TRUE, minuteMETsNarrow, by="Id")
n_distinct(dailyact_true_METs$Id)

# copy to work MAYBE CAN ELIMINATE AND RELOAD IF NEEDED
# make a copy to work with
copy_dailyActivity <- dailyActivity

# create a column for the total of minutes used
copy_dailyActivity <-    copy_dailyActivity %>% 
  mutate(total = VeryActiveMinutes + FairlyActiveMinutes +
                        LightlyActiveMinutes + SedentaryMinutes)
# create a column "valido" boolean to check if total is 1440 minutes or not
copy_dailyActivity <- copy_dailyActivity %>% 
  mutate(valido = total == 1440)
   
# Two data frames to facilitate work 
copy_dailyActivity_TRUE <- subset(copy_dailyActivity, valido == TRUE)
copy_dailyActivity_FALSE <- subset(copy_dailyActivity, valido == FALSE)



```
All graphs together

```{r graphs, echo=FALSE}
# graficar para ver tendencia
ggplot(data = metnarrow_calories) +
        geom_point(aes(x=METs, y=Calories), color = "blue", alpha = 0.5) +
        labs(title = "Relation between METs and Calories", x = "METs", y = "Calories")
# No hay relacion al menos visual

ggplot(data = day_steps_sleep) +
        geom_point(aes(x=StepTotal, y=TotalMinutesAsleep), color = "red", alpha = 0.5) +
        labs(title = "Relation between Steps and Sleep", x = "Steps", y = "Minutes_Sleep")

# plotting the full day group
ggplot(data = copy_dailyActivity_TRUE) +
        geom_point(aes(x=TotalSteps, y=Calories), color = "blue", alpha = 0.5) +
        labs(title = "Relation between Steps and Calories", x = "Steps", y = "Calories")

ggplot(data = copy_dailyActivity_FALSE) +
        geom_point(aes(x=TotalSteps, y=Calories), color = "red", alpha = 0.5) +
        labs(title = "Relation between Steps and Calories", x = "Steps", y = "Calories")

ggplot(data = dailyact_true_sleep) +
      geom_point(aes(x=TotalSteps, y=TotalMinutesAsleep), color = "purple", alpha = 0.5) +
        labs(title = "Relation between Steps and Sleep", x = "Steps", y = "Sleep in Minutes")

ggplot(data = dailyact_true_weight) +
      geom_point(aes(x=BMI, y=SedentaryMinutes), color = "orange", alpha = 0.5) +
        labs(title = "Relation between Sedentary and BMI", x = "BMI", y = "Sedentary  Minutes")

```

```{r correlation, echo=FALSE}
# computing the correlation for the two groups with Calories
cor(copy_dailyActivity_TRUE$TotalSteps, copy_dailyActivity_TRUE$Calories)

cor(copy_dailyActivity_FALSE$TotalSteps, copy_dailyActivity_FALSE$Calories)


# Correlation total steps vs heart rate
cor(dailyact_true_heart$TotalSteps, dailyact_true_heart$Value)
#  0.05 low correlation

# correlation total steps vs TotalMinutesAsleep
cor(dailyact_true_sleep$TotalSteps, dailyact_true_sleep$TotalMinutesAsleep)
# negative correlation  -0.2549887

# correlation total steps vs weightKg 
cor(dailyact_true_weight$TotalSteps, dailyact_true_weight$WeightKg)
# low correlation  0.1238845

# correlation total steps vs METs
cor(dailyact_true_METs$TotalSteps, dailyact_true_METs$METs)
# low correlation  0.1393237

# Correlation within dailyActivity total steps vs Calories
cor(dailyact_true_METs$TotalSteps, dailyact_true_METs$Calories)
# High correlation  0.7323676


cor(dailyact_true_weight$WeightKg, dailyact_true_weight$Calories)

```
* Summaries all together

```{r summary_dailyactivity, echo=FALSE}
# for dailyActivity
dailyActivity %>%  
  select(TotalSteps,
         TotalDistance,
         SedentaryMinutes) %>%
  summary()

# For the sleepDay dataframe:
sleepDay %>%  
  select(TotalSleepRecords,
  TotalMinutesAsleep,
  TotalTimeInBed) %>%
  summary()

```



*dailyActivity*
dailyActivity, check records that add to full day vs  records that do not<br>
How do the correlation with steps behave for the two groups<br>

Records that do not add to a full day, have a more disperse correlation between totalsteps and calories
The correlations values are 0.733 when the records have a full day
and 0.46 when the records do not have a full day


Merged dailyActitivity and the files heart rate, sleepDay and weightLogInfo to review correlations<br>
The group with full day.



Now you can explore some different relationships between activity and sleep as well. 
For example, do you think participants who sleep more also take more steps or fewer steps per day?
check chunk merge_3, negative correlation  -0.2549887
Check Graph and correlation in the next chunk

Is there a relationship at all? How could these answers help inform the marketing strategy of how you position this new product?

This is just one example of how to get started with this data - there are many other files and questions to explore as well!



